---
date: ' 2020-03-18T15:36:24 '
draft: false
linktitle: Lab 2
menu: 
     MES623Econ:
          parent: Lab Material
          weight: 2
title:   Data Types, Functions, and Loops
type: docs
weight: 2
output:
  blogdown::html_page:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="C:\\Users\\lxd312\\Documents\\GitHub\\academic-template\\static\\files")
```

## Data Types: Vectors and Matrices
### Vectors
Vectors are one-dimensional and can be created in many ways.
If you just want numbers from one to 10, create the object "vec" and use the colon ":" to indicate "to." This will automatically create one-digit increments.
If you want to control the increments, use the function seq(). The arguments are: from, to, and by.**Which argument controls the increment?**

```{r}
vec = 1:10
vec

vec_1 = seq(from=1,to=10,by=.5)
vec_1

```

What if we wanted the same number multiple times? That's easy--, we just use the rep() function. The basic arguments for this function are as follows: "x" (the object/number we are repeating),"times"", and "each". What if we want to repeat the number 1, 10 times?
```{r}
vec_2 = rep(x = 1,times = 10)
vec_2

```

Say we want to combine multiple vectors together. We use the "c()" function.
```{r} 
vec_12 = c(vec_1,vec_2)
vec_12 

# Alternatively
vec_12_alt = c(seq(from=1,to=10,by=.5),rep(1,10))
vec_12_alt
```

### **Exercise 1:**
Write the code to replicate the sequence 1:4 (by 1's), 3 times.
```{r, echo = FALSE, eval = TRUE}
vec_3 = rep(seq(1,4,1),times = 3)
vec_3
```

Want this to be ordered With three 1's, three 2's, etc.? Use the rep() function, the generic vector creation with a colon, and the argument "each = 3."
**Note:** We can also use these functions for character strings (ex: rep("dog", times = 4).
```{r}
vec_4 = rep(1:4, each = 3)
vec_4
```


You can also use basic mathematical operations on vectors but only if they are numeric and the same length. Uncomment the bad code to see the errors you may get.
```{r}
vec
vec_2
vec_add = vec+vec_2
vec_add

# This won't work ... why not?
# vec_bad = vec_2+vec_12
# vec_bad

# This won't work ... why not?
# vec_d = rep("dog", times = 4)
# vec_bad_1 = vec_d + vec_2
# vec_bad_1
```
**ProTip:** Since a vector has only one dimension to index it (determine what's in a specific position), you only need square brackets. Ex: vec_add[6] will return the value 7.

## Matrices
Matrices are multi-dimensional. We can create a matrix and put anything we want into it by using the matrix() function and it's basic arguments: data, nrow, and ncol. Let's put 1's in a 3 x 2 matrix.
```{r}
mat_1 = matrix(data =1, nrow = 3, ncol = 2)
mat_1 
```
**Note:** If you forget the "data =" in the function R will fill in your matrix with NAs. Try it for yourself.

We can add new info to our matrix, either by the columns using cbind() or by the row using rbind(). Here is an example of adding an additional row to our established matrix.
```{r}
mat_new = rbind(mat_1,rep(2,2))
mat_new
```

### **Exercise 2:**
We can find a specific entry in our matrix by indexing using square brackets, the row number, a comma, and then the column number. Let's figure out what is in the 4rd row and the 2nd column.
```{r, eval = FALSE, echo = FALSE}
mat_new[4,2]

mat_new

```



### Math with matrices
You can apply mathematical operations accross whole matrices. Example: Let's multiply each element by a scalar of  3.
```{r}
three_mat = mat_new * 3
three_mat
```

Let's kick it up a notch! What if we have two matrices that we want to multiply together? We can use the matrix operator " %*% ": Click the link to learn more about **[matrix alebra](https://www.statmethods.net/advstats/matrix.html)**.
```{r}

mat_2 = matrix(4,2,2)
mat_2

# remind us of what mat_1 was
mat_1

mat_12 = mat_1%*%mat_2
mat_12

# This won't work ... why not?
# three_mat%*%mat_1

```
**Note:** To multiply matrices make sure that the the number of columns in the 1st matrix equals the number of rows in the 2nd matrix. Use one of the functions we learned from last week to check the matrix dimension (hint, hint). 


Other data types such as lists and arrays exist. list() can have more than one type of data (i.e., numerics and character strings). Arrays have a dimension of 3 but all layers have to have the same dimensions.
```{r}
# Creating a List
pop_parms_list <- list("r" = 2.5, "dependent" = TRUE, "init.pop" = seq(0,1000,100))
pop_parms_list

# Creating an array
pop1 = matrix(1:6, nrow = 3 , ncol = 2) # matrix 1 to put in an array
pop2 = matrix(1:6, nrow = 3 , ncol = 2) # matrix 2 to put in an array
pops_array = array(c(pop1,pop2),dim = c(3,2,2)) # 3 rows, 2 columns, 2 entries
pops_array 
```



# Functions
Just like matrix(), header(), summary(), dim(), etc. are all functions, we can make our own functions.
Why do we make functions?

1.There is no available function to do what you want to do.

2.Work smarter -- not harder! You have a task that needs to be repeated, and you do not want to keep writing the same lines of code over and over again.


Functions must have a name and arguments, and they must return something. Arguments are the inputs you plan to use in your function. We often use the function return() to have our functions "print" out to the console or pass something on to another function. Example layout:
```{r, eval = FALSE}
func.name = function(arg1,arg2) 
  { # bracket to start function
  
  # code to do something, here
  
  return (something)

  } # bracket to end function
```

Let's do a simple matrix multiplication function.
```{r}
math.func = function (matrix.name,scalar) # 2 arguments to do something with
  {  
  
  new.matrix = matrix.name * scalar # code to do something with those arguments
  
  return(new.matrix) # code to print something 
  
  }# end function
```

Let's test it out! We need to create a matrix and then tell our function what the name of the matrix is and what value the scaler should be.
```{r}
mat.use = matrix(3,4,5)
math.func(matrix.name = mat.use, scalar = 2)
```
Look! It returned our "new.matrix"

### Defining Function Defaults for Testing
One day you may have a really complex function and you just want to make sure it works, without always putting in arguments to run it. Well you can set defaults for your arguments when you create your functions. Then to  run it quickly all you need is func.name(), R will do the rest. Lets try it out with our previous function.
```{r}
math.func = function (matrix.name = matrix(3,4,5) ,scalar = 1) 
  { # 2 arguments to do something with ... notice I now have defaults in place  
  
  new.matrix = matrix.name * scalar # code to do something with those arguments
  
  return(new.matrix) # code to print something 
  
  }

# Now we can just test to see if it works without putting in arguments, 
# just by typing the function name and parentheses
math.func()
```

### **Exercise 3:** 

Let's create our own function to determine the final grade for a student in this class. Make the default values so that the final grade is 100. Remember to test your function.

Things to think about:

1. What are the arguments/inputs?

2. What is the task we want to accomplish or calculation to be made using these arguments?

3. What do we expect our function to return?

```{r, echo = FALSE,results='hide', eval=TRUE}
myGrades = function (group.1 = 100,group.2 = 100, group.3 = 100, group.4 = 100, midterm.1 = 100, midterm.2 = 100, final = 100) # arguments with defaults 
{
  
  grade = 0.05*(group.1 +group.2+group.3+group.4) + 0.25*(midterm.1 + midterm.2) + 0.3*final
  
  return(grade)
} # end mygrades function

# Test the function... we should get a grade of 100 given the defaults
 myGrades()
```



#  For Loops
There are various types of loops but "for" loops are most common. You use loops when you want to do something iteratively. For loops have the following structure:
```{r, eval = FALSE}
for (i in 0:5) # a variable and range of values it will take ... "x" is not a good variable 
  { # a bracket to start
   
  #some cacluation or function to do something
  
  } # a bracket to end the for loop
```
Here "i" sequentially becomes a number from 0 to 5 and is used in each iteration of the loop.

Let's try a simple example where we want to add i to 5. Our output should consist of a list of values. We'll use the print() function so we can see what the loop is doing. 
```{r}
for (i in 0:5) 
  {
  x = i + 5 # Do something to i, iteratively
  
  print(x) # see what happens if you take this out
  
  # I like to print what I am doing inside of a loop ... 
  # uncomment the next line to see what it does
  
  # cat(paste("i= ", i, "and i + 5 =", x,"\n")) 
  
  } # end for loop
```

Lets tell R to do something different in the loop if i = 3. We do this by using an "if" statement which can be interpreted as "if something is true, do this." Because this is a logical statement, we must use logical operators in our if statment (i.e. "==", "!="). We can also use "else" to succinctly tell R how to continue but this is not always necessary.
```{r}

for (i in 0:5) 
  {
 
  x = i + 5
  
  # but ...
  if(i == 3) {x = "NULL"}  # our new if statement
  
  cat(paste("i= ", i, "and i + 5 =", x,"\n")) 

  } # end for loop

# Alternatively
for (i in 0:5) 
  {
  
  if(i == 3) {x = "NULL"} else {x = i + 5} # our new if else statement
  
  cat(paste("i= ", i, "and i + 5 =", x,"\n")) 
  
  } # end for loop
```


Sometimes we only want to go to a certain point in the loop and then tell R to break out of the loop. Let's tell R to breakout of the loop when x = 7 using the command "break."
```{r}

for (i in 0:5) 
  {
  
  x = i + 5 # Do something to i, iteratively
  
  if(x > 7) {break} # Here is that "break" command
  
  cat(paste("i= ", i, "and i + 5 =", x,"\n")) 
  
  } # end loop
```

# **All Together**


### **Exercise 4:** 

Read in the csv file "StudentGrades". Find the final grade for each student using our mygrades function and a loop. Hint: Some indexing will need to take place.Print out a statement detailing each student's final class grade using a combination of the cat() and paste() functions. Here is an example output.

```{r, eval = FALSE, echo = FALSE}
# Create the data
st.1 =sample(seq(40,100,by=10), 7, replace = TRUE)
st.2 =sample(seq(40,100,by=20), 7, replace = TRUE)
st.3 =sample(seq(40,100,by=5), 7, replace = TRUE)
st.4 =sample(seq(40,100,by=10), 7, replace = TRUE)
st.5 =sample(seq(40,100,by=20), 7, replace = TRUE)
st.6 =sample(seq(40,100,by=5), 7, replace = TRUE)
st.7 =sample(seq(40,100,by=10), 7, replace = TRUE)
st.8 =sample(seq(40,100,by=20), 7, replace = TRUE)

all_grades = as.matrix(rbind(st.1,st.2,st.3,st.4,st.5,st.6,st.7,st.8))

colnames(all_grades) = c("group.1","group.2","group.3","group.4", "mid.1", "mid.2", "final")

write.csv(all_grades, "StudentGrades.csv")
```
```{r, eval = TRUE, echo = FALSE}
data = read.csv("MES623_Lab2_StudentGrades_use.csv")

for (i in 1:8){
 
 # i = 1
  
  score =  myGrades(group.1 = data[i,2], group.2 = data[i,3], group.3 = data[i,4], 
                    group.4 = data[i,5], midterm.1 = data[i,6], midterm.2 = data[i,7], 
                   final=data[i,8]) # Index the matrix for each student and each grade that is needed


  
  cat(paste("Student", i, "has a grade of ", score,"\n"))
}


```


# Additional Resources
[Ed Rubin's Econometrics Course Notes](http://edrub.in/ARE212/resources.html)

[Quick-R By DataCamp](https://www.statmethods.net/)

Google!
